*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-10 Code/Manuscript_O2.Rmd
Fixed italics for genus Synechocystis

*Prochlorococcus* genomes encode 4 FtsH proteins [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], henceforth referred to
as FtsH1-4, homologs to the characterized FtsH isoforms of the model
freshwater cyanobacterium *Synechocystis sp. PCC6803*, and with presumably
parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher
light, HLI MED4 upregulates expression of FtsH1 and FtsH2
[@bonisteelStrainSpecificDifferences2018], homologs to the
*Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012]. In
contrast, representative LLIV strain MIT9313 shows no induction of
expression of these FtsH protease isoforms when shifted to high light,
and thus has fewer of these FtsH hexamers serving each photosystem [@bonisteelStrainSpecificDifferences2018].
Transcript analysis demonstrates that MIT9313 expressed primarily FtsH3,
homologous to *Synechocystis* *sll*1463, possibly involved in PSI
biogenesis [@mannInvolvementFtsHHomologue2000; @krynickaFtsH4ProteaseControls2022,
@konikCyanobacterialFtsH4Protease2024]. FtsH3 expression did not
increase in response to light stress in MIT9313
[@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady
low light, clade LLIV *Prochlorococcus* instead allocate resources to
processes other than dynamic regulation of PSII repair.

@@

FtsHTable <- tibble(Organism = c("*Prochlorococcus marinus*", "*Synechocystis sp. PCC6803*", "*Synechocystis sp. PCC6803* isoform", "Function"),
                    FtsH1 = c("FtsH1", "*Slr*0228", "FtsH2", "PSII Repair"),
                    FtsH2 = c("FtsH2", "*Slr*1604", "FtsH3", "PSII Repair"),
                    FtsH3 = c("FtsH3", "*Sll*1463", "FtsH4", "PSI biogenesis"),
                    FtsH4 = c("FtsH4", "*Slr*1390", "FtsH1", "Cell viability")
                    )

>>>>

*Prochlorococcus* genomes encode 4 FtsH proteins [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], henceforth referred to
as FtsH1-4, homologs to the characterized FtsH isoforms of the model
freshwater cyanobacterium *Synechocystis sp.* PCC6803, and with presumably
parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher
light, HLI MED4 upregulates expression of FtsH1 and FtsH2
[@bonisteelStrainSpecificDifferences2018], homologs to the
*Synechocystis* *slr*0228 and *slr*1604, implicated in PSII repair [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012]. In
contrast, representative LLIV strain MIT9313 shows no induction of
expression of these FtsH protease isoforms when shifted to high light,
and thus has fewer of these FtsH hexamers serving each photosystem [@bonisteelStrainSpecificDifferences2018].
Transcript analysis demonstrates that MIT9313 expressed primarily FtsH3,
homologous to *Synechocystis* *sll*1463, possibly involved in PSI
biogenesis [@mannInvolvementFtsHHomologue2000; @krynickaFtsH4ProteaseControls2022,
@konikCyanobacterialFtsH4Protease2024]. FtsH3 expression did not
increase in response to light stress in MIT9313
[@bonisteelStrainSpecificDifferences2018]. Through adaptation to steady
low light, clade LLIV *Prochlorococcus* instead allocate resources to
processes other than dynamic regulation of PSII repair.

@@

FtsHTable <- tibble(Organism = c("*Prochlorococcus marinus*", "*Synechocystis sp.* PCC6803", "*Synechocystis sp.* PCC6803 isoform", "Function"),
                    FtsH1 = c("FtsH1", "*Slr*0228", "FtsH2", "PSII Repair"),
                    FtsH2 = c("FtsH2", "*Slr*1604", "FtsH3", "PSII Repair"),
                    FtsH3 = c("FtsH3", "*Sll*1463", "FtsH4", "PSI biogenesis"),
                    FtsH4 = c("FtsH4", "*Slr*1390", "FtsH1", "Cell viability")
                    )
